The Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-04-02 paper.tex
section 3 finished

However, we think that this is not the full story. When agreeing to understand anthropomorphism as a specific type of \textit{experience} that arises in an
\textit{interaction} between a set of user expectations and the external reality, 
%(see Figure~\ref{fig:anthropomorphism_and_interaction})
it becomes clear that this interaction happens in a specific context, with a specific history. We think that the context and the history of the interaction (in Figure \ref{fig:factors} denoted as \textit{time}), are two more factors that determine anthropomorphism.

Consequently, we propose a holistic understanding of anthropomorphism as a phenomenon which is determined by multiple factors:

\item ~the \textbf{characteristics of the robot's design} (degree of anthropomorphic form),
\item ~the \textbf{psychological determinants in the human user} (see
\cite{epley_seeing_2007}),
\item ~the \textbf{characteristics of the context} (situation) in which the interaction occurs, and
\item ~the fact of the user getting \textbf{familiar with the robot over time} (getting used to a system the more and longer one interacts with it)

Further, we propose that
anthropomorphism is a multi-layered phenomenon which arises in different levels \citep{persson_anthropomorphism_2000}, and which is likely to evolve with continued interaction.

@@

Second, we would like to take up the proposition already made by other researchers (\eg \cite{persson_anthropomorphism_2000,takayama_perspectives_2012,ruijten_introducing_2014}) to not viewing anthropomorphism as being either there or not there, but to discriminate different levels of
anthropomorphism (low, and high anthropomorphism, for instance).
This seems to make sense, since each of the four proposed factors has its own characteristics and involves specific
types of user expectations, for instance. Such a differentiation in the understanding of anthropomorphism could further allow to draw inferences
on the human-robot relationship based on the observed level of anthropomorphism.

To facilitate things, we introduce the notation \Ant[R,H,C,t] to reflect these multiple factors: we suggest here that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build
an interaction context $C = \{s_0,...,s_n\}$, and the duration of the
interaction itself, $t$.

@@

With having both before-mentioned aspects in mind (multi-factors that determine anthropomorphism and multi-levels of the observable phenomenon), it becomes clear that anthropomorphism cannot be something fixed, that once observed in a human-robot interaction always remains the same. We think, that anthropomorphism evolves and shows dynamics. This idea is what we would like to formalize in the next section.

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Though conflicts were
found\fxwarning{SL: which conflicts are we talking about?}, the HRI
community more or less accepted viewing anthropomorphism as a static feature
that once observed during a short-term interaction reflects a sustaining social
effect.

\citet{kanda_interactive_2004} point out, however, that relationships among
people evolve over time and, in a similar manner, one can expect people's attitude toward
technological artifacts and their relationship with them to evolve as well.
In their study, \citet{kanda_interactive_2004} were also among the first to
describe empirically \emph{novelty effects} during a long-term study with an
interactive robot in a school.

From a theoretical perspective, we also expect anthropomorphism to change over
time, with growing experience with a robot. As discussed in the previous
sections, anthropomorphism means in essence \emph{reasoning about and perceiving
something non-human (and unknown or unfamiliar) based on one's
representation of the familiar and well-known concept of being human (or
one-self)}. According to \citet{epley_when_2008}, the basic operations that
underlie these inductive inferences are the \emph{acquisition of knowledge}, the
\emph{activation or elicitation of knowledge}, and the \emph{application of
activated knowledge at the time of the judgment}. By applying knowledge to a
given situation or interaction, it is implied that one attempts to correct,
adjust, or maybe integrate less accessible information into a more automatically
activated default representation. Epley argues that this process can be seen as
a process of correction, and thus, it is often insufficient leaving final
judgments biased in the direction of the initially activated representation.
Refinements to the initial representation mean something evolves. This can be
due to the fact that a person's knowledge base changes constantly with newly
acquired things, or growing experiences with a robot. Consequently, it is likely
that the ``need to anthropomorphize'' a robot decreases over time (from the
human-centered point of view), since the user's knowledge about the robot
evolves, and also the user familiarizes herself with the robot. Consequently,
the robot should become more predictable and more familiar to the human user and
inferences about it can be made based on the acquired knowledge.\fxwarning{SL:
Rephrase the last part of this paragraph}

In other words, the argumentation is that humans have a motivation to interact
effectively in their environment \textit{(effectance)} \citep{epley_when_2008}.
This means that humans are motivated to be able to understand and predict the
environment and agents around them. Through this process, uncertainty is
reduced, and consequently, a first unfamiliar robot, can be explained with means
that are different from anthropomorphism.\fxwarning{SL: Paragraph unclear to me.
Rephrase it?}

In the introduction, we reported on a few selected experimental results that
illustrate how the user tendency to anthropomorphize decreased over time.
Combined with the extensive discussion on the nature and factors of
anthropomorphism that we proposed at section~\ref{sec:anthropomorphism}, we
introduced in the following section a novel qualitative model of the evolution
of anthropomorphism over time, \ie a model of the dynamics of anthropomorphism.

>>>>

However, we think that this is not the full story. When agreeing to understand anthropomorphism as a specific type of \textit{experience} that arises in an
\textit{interaction} between a set of user expectations and the external reality \citep{persson_anthropomorphism_2000}, 
%(see Figure~\ref{fig:anthropomorphism_and_interaction})
it becomes clear that this interaction happens in a specific context, with a specific history. We think that the context and the history of the interaction (in Figure \ref{fig:factors} denoted as \textit{time}), are two more factors that determine anthropomorphism.

Consequently, we propose a holistic understanding of anthropomorphism as a phenomenon which is determined by multiple factors (see also Figure \ref{fig:factors}):

\begin{enumerate}
\item ~the characteristics of the \textbf{robot's design} (degree of anthropomorphic form),
\item ~the psychological determinants in the \textbf{human user} (see
\cite{epley_seeing_2007}),
\item ~the characteristics of the \textbf{context / situation} in which the interaction occurs, and
\item ~the fact of the user getting familiar with the robot \textbf{over time} (duration of interaction)
\end{enumerate}

@@

When taking into account these multiple factors, a logical consequence seems to be that anthropomorphism is not either there or not there but that it arises in different levels, may they be distinct, having multiple layers \citep{persson_anthropomorphism_2000}. or a continuum being describes in low and high anthropomorphism \citep{ruijten_introducing_2014}. For instance,

Moreover, we propose that anthropomorphism shows not only variances determined by these four main factors but that -- the fourth factor implies this -- anthropomorphism is likely to evolve over time with continued interaction between the human and the non-human agent.

To facilitate things, we introduce the notation \Ant[R,H,C,t] to bring these multiple factors together. To take into account dynamics over time, we suggest that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build
an interaction context $C = \{s_0,...,s_n\}$, and the duration / history of the
interaction itself, $t$.

@@

@@

So far, the HRI community has not much investigated how anthropomorphism in
human-robot interactions evolves over time (during the process of
\emph{adopting} a robot, for instance). Though conflicts were
found in the understanding of what anthropomorphism is, the HRI
community more or less accepted viewing anthropomorphism as a static feature
that once observed during a short-term interaction reflects a sustaining social
effect. However, if we agree that anthropomorphism arises in an interaction, and thus reflects a specific relation, we cannot accept it as a static phenomenon.

Similarly, \citet{kanda_interactive_2004} point out that relationships among
people evolve over time and, in a similar manner, one can expect people's attitude toward
technological artifacts and their relationship with them to evolve as well.
In their study, \citet{kanda_interactive_2004} were among the first to
describe empirically \emph{novelty effects} during a long-term study with an
interactive robot in a school.

From a theoretical perspective, we also expect anthropomorphism to change over
time, with growing experience with a robot. As discussed in the previous
sections, anthropomorphism means in essence \emph{reasoning about and perceiving
something non-human (and unknown or unfamiliar) based on one's
representation of the familiar and well-known concept of being human (or
one-self)}. According to \citet{epley_when_2008}, the basic operations that
underlie these inductive inferences are the \emph{acquisition of knowledge}, the
\emph{activation or elicitation of knowledge}, and the \emph{application of
activated knowledge at the time of the judgment}. By applying knowledge to a
given situation or interaction, one also attempts to correct,
adjust, or maybe integrate less accessible information into a more automatically
activated default representation.
%Epley argues that this process can be seen as
%a process of correction, and thus, it is often insufficient leaving final
%judgments biased in the direction of the initially activated representation.
%Refinements to the initial representation mean something evolves.
This process of refining the applied knowledge are natural because a person's knowledge base changes constantly with newly
acquired things. For instance, a person can acquire new knowledge about a robot when interacting with it. With continued interaction with the unknown system also the user's expertise with the system grows, and it can be expected that in turn, the user will become able to better predict the previously unknown system. Consequently, as the robot becomes more predictable, there is less uncertainty, and the ``need to anthropomorphize'' the robot decreases (from the
human-centered point of view).

In other words, humans have a motivation to interact
effectively in their environment \textit{(effectance)} \citep{epley_when_2008}.
This means that humans are motivated to be able to understand and predict the
environment and the agents around them. Thus, a person will seek to become familiar with what is unfamiliar. Anthropomorphism is encouraged when an agent is unfamiliar. Consequently, in the process of familiarizing oneself with an agent, the tendency to anthropomorphize this agent should decrease.

We can not be sure about how exactly the process of familiarization to an unfamiliar agent (based on continued interaction with the robot) looks like. In turn, we can not make a precise statement about how continued interaction decreases anthropomorphism but we suspect there is a link between the two.

In the introduction, we reported on a few selected experimental results that
illustrate how the user tendency to anthropomorphize decreased over time.
Combined with the extensive discussion on the nature and factors of
anthropomorphism that we proposed at section~\ref{sec:anthropomorphism}, we
introduce in the following section a novel qualitative model of the evolution
of anthropomorphism over time, \ie a model of the \emph{dynamics of anthropomorphism}.
